---
title: "Project 1: Data Analysis"
author: "Ellen Kim"
date: "`r Sys.Date()`"
 
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library (readr)
library(dplyr)
library(stringr)
library (readr)

urlfile="https://raw.githubusercontent.com/Lnkiim/DATA607FALL202/master/Project1/tournamentinfo.txt"
 
 
```

### Read text file into dataframe

```{r code-chunk-label}

# reads in data, using tab as delimiter
raw_tournament_df <- read.table(urlfile, header = FALSE, sep = "\t")

# creating new columns by parsing by |
tournament_df2 <- raw_tournament_df %>% separate(V1, c("Pair","player_name","total_pts","round1", "round2", "round3", "round4","round5", "round6","round7"), sep ="(\\|)")

# exclude all rows with --- in the 'Pair' column
tournament_df2 <- dplyr::filter(tournament_df2, !grepl("-------",Pair))

tournament_df3 <- tournament_df2

# Regex function that can be used to extract all numerical characters
number_extract <- function(string){ 
  str_extract(string, "\\d+\\.*\\d*")
}

# Regex function that can be used to extract all alphabet characters
string_extract <- function(string){ 
  str_extract(string, "\\D+\\.*\\D*")
}
 
# Extract opponent's playerId
# only populate id if its number type
tournament_df3$playerId <- number_extract(tournament_df3$Pair) 

# apply function above
for (column_name in c("Pair", "round1", "round2", "round3", "round4","round5", "round6","round7")){ 
  number_extract(column_name) 
}

#only populate if alphabet type
tournament_df3$city <- string_extract(tournament_df3$Pair) 

# every other row does not have tie to player so we cascade playerId down using a window function
tournament_df3 <- 
  tournament_df3 %>% 
    fill(playerId) %>% 
    mutate(
      w = ifelse(cummax(playerId) > 2, cummax(playerId), playerId)
    )
 
# separate odd and even numbered rows to be in separate dataframes for data cleaning
tournament_df3 <- tournament_df3 %>% 
  group_by(playerId) %>%
  mutate(ranking = rank(row_number()) )

# create 2 separate dataframes 
rounds_df1 = tournament_df3 %>% filter(ranking  == 1)
user_points_df1 = tournament_df3 %>% filter(ranking  == 2)
```

### Dataframe that contains information per round
Extract all opponent numbers

``` {r}
 
# extract opponent's playerId for each round
rounds_df1$opponent1 <- number_extract(rounds_df1$round1) 
rounds_df1$opponent2 <- number_extract(rounds_df1$round2) 
rounds_df1$opponent3 <- number_extract(rounds_df1$round3) 
rounds_df1$opponent4 <- number_extract(rounds_df1$round4) 
rounds_df1$opponent5 <- number_extract(rounds_df1$round5) 
rounds_df1$opponent6 <- number_extract(rounds_df1$round6) 
rounds_df1$opponent7 <- number_extract(rounds_df1$round7) 
```


### Dataframe that contains user-level data
```{r}
# extract the pre-score

# Regex function that can be used to extract all numerical characters
pre_score <- function(string){ 
  str_extract(str_extract(string, "R:\\s+\\d+[a-zA-Z ]"), "\\d+")
}

# Applying function to extract preScore
user_points_df1$preScore <- pre_score(user_points_df1$player_name) 

# New df with subset of columns: state, playerId, prescore
user_points_subset_columns <- c("Pair", "playerId","preScore")
user_points_subset <-user_points_df1[user_points_subset_columns]

# Rename user_state
colnames(user_points_subset)[1] <- c("user_state")
```

### How many games did each player play?
We only count W/L/D as games played and we need count of games played to be denominator for the ELO calculation.

```{r}
# creating dataframe which will be filled later
rounds_df1$gamesPlayed <- NA

# function to count number of games with that are either W,L, or D.
count_games <- function(string){ 
  if (grepl("[WLD]", string) == TRUE) 1 else 0
}

# iterate through rows of the df to apply function, then sum, insert value into new column
for(i in 1:nrow(rounds_df1)) {
    row <- rounds_df1[i,] 
    rounds_df1$gamesPlayed[i] <-  count_games(row$round1) + count_games(row$round2) + count_games(row$round3) + count_games(row$round4) + count_games(row$round5) + count_games(row$round6) + count_games(row$round7)
}

# New df with subset of columns
rounds_subset_cols <- c("playerId", "player_name", "total_pts", "opponent1","opponent2","opponent3","opponent4","opponent5","opponent6","opponent7", "gamesPlayed" )
rounds_subset <- rounds_df1[rounds_subset_cols]

```

 
### Get preScore of Opponents
```{r}

# merge the two existing dataframes
masterdf_1 <- merge(x = rounds_subset, y = user_points_subset, by = "playerId", all = TRUE)


#create new column for sum of all opponent's scores
masterdf_1$sumOppPreScore <- NA

# Change datatype from char to numeric
cols_num <- c("playerId","total_pts", "opponent1","opponent2","opponent3","opponent4","opponent5", "opponent6" ,"opponent7", "preScore", "sumOppPreScore")
masterdf_1[cols_num] <- sapply(masterdf_1[cols_num],as.numeric)
sapply(masterdf_1, class)
 
 
#handle NA's where player didn't have an opponent on any given week
clear_score <- function(score_opp){
  if(is.na(score_opp)) 0 else score_opp  
}

# sum all the opponent's preScores
for(i in 1:nrow(masterdf_1)){
  opp1_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent1"] ][1])
  opp2_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent2"] ][1])
  opp3_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent3"] ][1])
  opp4_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent4"] ][1])
  opp5_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent5"] ][1])
  opp6_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent6"] ][1])
  opp7_score <- clear_score(masterdf_1$preScore[masterdf_1$playerId == masterdf_1[i,"opponent7"] ][1])
  tot_opp_score <-  opp1_score + opp2_score +  opp3_score + opp4_score +  opp5_score + opp6_score + opp7_score
  masterdf_1$sumOppPreScore[i] <- tot_opp_score[1]
}


```


### Calculate Average Opponent's PreScore
Sum of opponents prescore / total games played
```{r}

masterdf_1$avg_opp_preScore <- masterdf_1$sumOppPreScore / masterdf_1$gamesPlayed
```


### Final Cleaning - subset
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
```{r}
final_subset <- c("player_name","user_state","total_pts", "preScore","avg_opp_preScore")
final_chess_data <- masterdf_1[final_subset]

#drop null row that used to be header
final_chess_data <- na.omit(final_chess_data)
```
 
 
### Write final dataframe to csv file
```{r}
write.csv(final_chess_data,"/Users/ellenkim/CUNY/data_acquisition_management/AcqMgmtRepot_DATA607FALL202/Project1/project1output.csv", row.names = TRUE)
```
 