---
title: "Week 10 Assignment – Sentiment Analysis"
author: "Ellen Kim"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)
library(tidytext)
library(textdata)
library(tidyr)
library(janeaustenr) 
library(ggplot2)
library(stringr)
```

### Re-create base analysis

Insert any text here.

```{r}

get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")

# tokenize lines to words
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
      ignore_case = TRUE
    )))
  ) %>%
  ungroup() %>%
  unnest_tokens(word, text)

# get joy words from nrc
nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")

# inner join "joy' sentiment words from Emma
tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)


# for every chunk of 80 lines, sentiment = pos - negative
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  # breaking lines of a book into chunks of 80
  count(book, index = linenumber %/% 80, sentiment) %>%
  # transposing sentiment values as columns
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
 


```



```{r}
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")

```


```{r}
# 2.3 Comparing the three sentiment dictionaries

library(tidyr)

pride_prejudice <- tidy_books %>%
  filter(book == "Pride & Prejudice")

# sentiment by 80 line chunks using AFINN
afinn <- pride_prejudice %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = linenumber %/% 80) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

# stack after getting sentiments of 2 diff lexicon dictionaries
# then calculate sentiment, partitioned by every 80 lines
bing_and_nrc <- bind_rows(
  # get sentiment for bing
  pride_prejudice %>%
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  # get sentiment for nrc
  pride_prejudice %>%
    inner_join(get_sentiments("nrc") %>%
      filter(sentiment %in% c(
        "positive",
        "negative"
      ))) %>%
    mutate(method = "NRC")
) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)




```




```{r}
# stack 3rd lexicon library AFINN and plot
bind_rows(
  afinn,
  bing_and_nrc
) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")

```





```{r}

get_sentiments("nrc") %>%
  filter(sentiment %in% c(
    "positive",
    "negative"
  )) %>%
  count(sentiment)

get_sentiments("bing") %>%
  count(sentiment)
```




```{r}
#2.4 Most common positive and negative words

bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

# most common words
bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(
    y = "Contribution to sentiment",
    x = NULL
  ) +
  coord_flip()

```



```{r}
# creating custom stop words list by appending to stop_words
custom_stop_words <- bind_rows(
  tibble(
    word = c("miss"),
    lexicon = c("custom")
  ),
  stop_words
)

library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

```


```{r}
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(
    colors = c("gray20", "gray80"),
    max.words = 100
  )
```




```{r}
# partitioned by sentence. delimter is period.
PandP_sentences <- tibble(text = prideprejudice) %>%
  unnest_tokens(sentence, text, token = "sentences")

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text,
    token = "regex",
    pattern = "Chapter|CHAPTER [\\dIVXLC]"
  ) %>%
  ungroup()

austen_chapters %>%
  group_by(book) %>%
  summarise(chapters = n())


bingnegative <- get_sentiments("bing") %>%
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords / words) %>%
  filter(chapter != 0) %>%
  top_n(1) %>%
  ungroup()

```

### My Corpus
My corporus is a dataset which has reviews for gourmet foods sold on Amazon. I would like to compare the sentiment analysis results with the stars given for each product. 

```{r}

library(readr)

# file is too large. won't upload to git with free account
food_reviews <- read_csv(file= "Reviews.csv")


# data is taking too long to render
# Lets partition the data by random sampling 

## 75% of the sample size
smp_size <- floor(0.50 * nrow(food_reviews))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(food_reviews)), size = smp_size)

# creates smaller set but also create identity column
train <- food_reviews[train_ind, ]
#test <- food_reviews[-train_ind, ]

 
tidy_reviews <- train %>%
  select(Id, ProductId, UserId, Score, Text) %>% 
    unnest_tokens(word, Text) %>%
    anti_join(stop_words)

tidy_reviews_bing <- tidy_reviews %>%
  inner_join(get_sentiments("bing")) %>%
    count(ProductId, Id, sentiment, Score) %>%
      spread(sentiment, n, fill=0) %>%
      mutate(sentiment = positive - negative)
  
tidy_reviews_bing %>%
  group_by(Score) %>%
    summarize(avg_score = mean(sentiment))
 
```

### Conclusion
We can see that the average score from the sentiment scoring is aligned with the score given by the reviewer. While the amazon reviews are on a scale of 1 to 5, the sentment count is not bound by the same scale. However, the average sentiment score is relative and does show that the lowest scored items (equal to 1) also has the lowest sentiment score, while the highest review score of 5, also has the highest sentiment scoring.
 
 
### A Different Lexicon
Previously we were using the Bing lexicon. Lets see if the "AFINN" lexicon gives us significantly different results. 
Methodology: Using tokenized dataframe, inner join with afinn lexicon. This lexicon outputs values from -5 to 5. The count by ProductId, Id, and Score, which results in a frequency for each sentiment value that occurs for one review. Next, create new feature called sent_times_n, which creates a score that takes frequency of each sentiment value into account. The final sentiment score, is the average score for a single review.

```{r}
tidy_reviews_afinn <- tidy_reviews %>%
  inner_join(get_sentiments("afinn")) %>%
    count(ProductId, Id, value, Score) %>%
      mutate(sent_times_n = value * n) %>%
      group_by(ProductId, Id,  Score) %>%
      mutate(sentiment_score = mean(sent_times_n)) %>%
      ungroup() 

tidy_reviews_afinn %>%
  group_by(Score) %>%
    summarize(avg_score = mean(sentiment_score))
```

#### Findings of Second Lexicon

The use of different lexicon again, align with the ratings, as well as the conclusions of the first lexicon. Relative to each other, the lowest rated items have the lowest sentiment score and the highest rated items have the highest sentiment scores. Because my analysis had was based on relativity, it doesn't make a signficant difference in which lexicon is used. 


```{r}

```










