---
title: "Assignment 3: - Character manipulation & Data processing "
author: "Ellen Kim"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r code-chunk-label}
library (readr)
urlfile="https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
college_majors <- read_csv(url(urlfile))

majors <- college_majors$Major

dataOrStatistics <- grep(pattern = "DATA|STATISTICS", majors, value = TRUE, ignore.case = TRUE)
```

### Exercise 2 
Write code that transforms the data below

```{r}
# Here we have some blob that we want to convert into a neat vector

some_str_blob <- '“blueberry”  “cantaloupe” “chili pepper” “cloudberry” “elderberry”   “lime”   “lychee”   “mulberry"   “olive”   “salal berry” '

words <- as.list(strsplit(some_str_blob, '\\s+')[[1]])

 
```
 
 
### Exercise 3
Describe, in words, what these expressions will match:
(assuming this is for regex in R not standard regex)
 

```{r}

library(stringr)
test <- list("777", "data\1\1", "anna", "2002", '"elle"', '"anna"', '"xoxox"', '"abc2342342343cba"')
str_view(test, '(.)\1\1', match = TRUE)
str_view(test, '"(.)(.)\\2\\1"', match = TRUE)
str_view(test, '(..)\1', match = TRUE)
str_view(test, '"(.).\\1.\\1"', match = TRUE)
str_view(test, '"(.)(.)(.).*\\3\\2\\1"', match = TRUE)
``` 
1) (.)\1\1
This will capture any non newline character, followed by '\1\1'.

2) "(.)(.)\\2\\1"
Matches character ", follwed by any 2 characters that are non-line break characters, followed by character in group 2, followed by character in group 1, followed by ". This would find 4 character palidromes that are enclosed in quotes.
 
3) (..)\1
Matches any 2 non-line break characters followed by \1 .

4) "(.).\\1.\\1"
This would find strings where every other character is the same three times, enclosed in quotes.

5) "(.)(.)(.).*\\3\\2\\1"
This would find a word that starts with three letters and ends with the same three letters in reverse, enclosed in quotes.


### Exercise 4
Construct regular expressions to match words that:

1) Start and end with the same character.
ANSWER = (.).*\1
 

2) Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
ANSWER = (..).*\1

3) Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
ANSWER = (.).*\1.*\1
  

```{r}

library(stringr)
test <- list("777", "data\1\1", "anna", "2002", '"elle"', '"anna"', '"xoxox"', '"abc2342342343cba"')
str_view(test, '(.).*\\1', match = TRUE)
str_view(test, '(..).*\\1', match = TRUE)
str_view(test, '(.).*\\1.*\\1', match = TRUE) 
``` 

